---
title: "Sample codes in lab package paper"
author: Yi-Ju Tseng, Chun Ju Chen
output: github_document
---

*lab*: An R package for generating analysis-ready data from laboratory records

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r install, eval=F}
remotes::install_github("DHLab-TSENG/lab")
```

```{r lib}
library(lab)
```

# Data Input - Sample data


```{r code}
head(labSample)
```


# LONIC Mapping

## Mapping Table

```{r map}
head(mapSample)
```

## LONIC Mapping

```{r LONIC}
loincSample <- mapLOINC(labData = labSample, labItemColName = ITEMID, mappingTable = mapSample)
loincSample[loincSample$SUBJECT_ID==36&loincSample$CHARTTIME=="2131-05-17",]
```

## Normal or Abnormal Test Results Identificaiton

```{r Ab}
loincMarkedSample <- getAbnormalMark(labData = loincSample,
                                     idColName = SUBJECT_ID,
                                     labItemColName = LOINC,
                                     valueColName = VALUENUM,
                                     genderColName = GENDER,
                                     genderTable = patientSample,
                                     referenceTable = refLOINC)
head(loincMarkedSample)

```

## Finding Related Labortory Tests

```{r find}
caseCreatinine <- searchCasesByLOINC(labData = loincSample,
                                     idColName = SUBJECT_ID,
                                     loincColName = LOINC,
                                     dateColName = CHARTTIME,
                                     condition = "Creatinine",
                                     isSummary = TRUE)

head(caseCreatinine)


```

# Time Series Analysis

## Deciding Width of Windows for Slicing Data

```{r window}
windowProportion <- plotWindowProportion(labData = loincSample,
                     idColName = SUBJECT_ID,
                     labItemColName = LOINC,
                     dateColName = CHARTTIME,
                     indexDate = first,
                     gapDate = c(30, 90, 180, 360),
                     topN = 5)

head(windowProportion$missingData)
print(windowProportion$graph)
```

## Slicing Data into Time-series Windows

```{r slice}
timeSeriesData <- getTimeSeriesLab(labData = loincSample,
                                   idColName = SUBJECT_ID,
                                   labItemColName = LOINC + LABEL,
                                   dateColName = CHARTTIME,
                                   valueColName = VALUENUM,
                                   indexDate = first,
                                   gapDate = 30,
                                   completeWindows = TRUE)
timeSeriesData[timeSeriesData$ID==36&timeSeriesData$LOINC=="2160-0"]
```


## Data Visualuzation
```{r vis}
timeSeriesPlot <- plotTimeSeriesLab(labData = timeSeriesData,
                                    idColName = ID,
                                    labItemColName = LOINC + LABEL,
                                    timeMarkColName = Window,
                                    valueColName = Nearest,
                                    timeStart = 1,
                                    timeEnd  = 5,
                                    abnormalMarkColName = NULL)

plot(timeSeriesPlot)
```
```{r vis2}
colnames(patientSample)[2]<-"ID"
timeSeriesDataMarked <- getAbnormalMark(labData = timeSeriesData,
                                     idColName = ID,
                                     labItemColName = LOINC,
                                     valueColName = Nearest,
                                     genderColName = GENDER,
                                     genderTable = patientSample,
                                     referenceTable = refLOINC)

timeSeriesPlotMarked <- plotTimeSeriesLab(labData = timeSeriesDataMarked,
                                    idColName = ID,
                                    labItemColName = LOINC + LABEL,
                                    timeMarkColName = Window,
                                    valueColName = Value,
                                    timeStart = 1,
                                    timeEnd  = 5,
                                    abnormalMarkColName = ABMark)

plot(timeSeriesPlotMarked)
```
# Data Imputation
```{r imp}
timeSeriesData[timeSeriesData$ID==36&timeSeriesData$LOINC=="2160-0"]

fullTimeSeriesData <- imputeTimeSeriesLab(labData = timeSeriesData,
                                   idColName = ID,
                                   labItemColName = LOINC + LABEL,
                                   windowColName = Window,
                                   valueColName = Mean & Nearest,
                                   impMethod = NOCB)

fullTimeSeriesData[timeSeriesData$ID==36&timeSeriesData$LOINC=="2160-0"]

```

# Analysis Ready Data Generation
```{r ARD}
wideTimeSeriesData <- wideTimeSeriesLab(labData = timeSeriesData,
                                        idColName = ID,
                                        labItemColName = LOINC + LABEL,
                                        windowColName = Window,
                                        valueColName = Nearest)
head(wideTimeSeriesData)

```

```{r ARDimp}
fullTimeSeriesData <- imputeTimeSeriesLab(labData = timeSeriesData,
                                   idColName = ID,
                                   labItemColName = LOINC + LABEL,
                                   windowColName = Window,
                                   valueColName = Nearest,
                                   impMethod = mean)

wideTimeSeriesData <- wideTimeSeriesLab(labData = fullTimeSeriesData,
                                        idColName = ID,
                                        labItemColName = LOINC + LABEL,
                                        windowColName = Window,
                                        valueColName = Nearest)
head(wideTimeSeriesData)

```